Synthesizing Symmetric Lenses

### 2018-07-10 papers/SymmetricLenses/main.tex
in progress edits to section 1

To overcome this second difficulty, we turn to information theory 
to characterize the likely and unlikely lenses. 
We consider a ``likely'' lens to be one that propagates a lot of
information from source to 
target and vice versa. Conversely, an ``unlikely'' lens requires a
large amount of additional information to recover the target given a
source and vice versa.  
This categorization follows from the intuition
that users will typically want to synchronise formats with lots
of shared information, and so lenses that preserve 
more information are preferable.

To implement this
intuition, we turn to \emph{stochastic regular expressions} (SREs). Stochastic
regular expressions define a set of strings, as do ordinary regular expressions,
but they also provide the expected distribution of strings within that set. Such
distributions allow us to compute the amount of information gained when
presented an element of the set.

\begin{figure}
  \includegraphics[width=.5\textwidth]{high-level-algorithm.pdf}
  \caption{Schematic diagram for our symmetric lens synthesis algorithm. Regular
    expressions \Regex and \RegexAlt, and a set of examples \Examples, are
    provided as input. \Expand searches through stochastic regular expression
    pairs, equivalent to the originals, and proposes them to \GreedySynth.
    \GreedySynth finds a lens, typed between the generated equivalent pairs.
    When the algorithm determines it has an optimal lens, it returns
  that lens. \ksf{Figure needs to be changed to call first search
  Expand; Are the S and T regular expressions (which should have bars)
  or stochastic regular expressions (which should be called such)?}}
  \label{fig:high-level-algorithm}
\end{figure}

With simple symmetric lenses and SREs in hand, we develop a new algorithm for
synthesizing likely lenses, based on the previous algorithm for synthesizing
bijective lenses. This algorithm is constructed from \Expand and \GreedySynth,
a pair of collaborating search procedures, shown in
Figure~\ref{fig:high-level-algorithm}. Given source and target SREs $S$ and $T$,
the \Expand searches for two \emph{related} SREs $S'$ and $T'$ that are
\emph{compatible} (a heuristic that estimates the likelihood our second
algorithm will succeed). $S'$ and $T'$ are related in the sense that they define
the same regular sets as $S$ and $T$ (respectively) and the distribution of
strings is identical. Indeed, we guarantee these two properties by implementing
a search procedure that obeys the star semi-ring axioms of regular expression
equivalence, and which we show does not alter the probability distributions of
the languages.

Given the two compatible SREs, \GreedySynth uses a greedy, type-directed
algorithm find a likely lens between them. If algorithm estimates that the lens
it has discovered is more likely than any future lens it can possibly find, the
search terminates. Otherwise, this lens is stored away (if it is the best found
so far) and the search continues with \Expand again.

>>>>

To overcome this second difficulty, we turn to information theory 
to characterize the likely and unlikely lenses. 
We consider a ``likely'' lens to be one that propagates a lot of
information from source to 
target and vice versa. Conversely, an ``unlikely'' lens requires a
large amount of additional information to recover the target given a
source and vice versa.  
This categorization follows from the intuition
that users will typically want to synchronise formats with lots
of shared information, and so lenses that preserve 
more information are preferable.
To implement this
intuition, we turn to \emph{stochastic regular expressions}
(SREs)~\cite{?}, which simultaneously define a set of strings \`a la
regular expressions and a probability distribution over those
strings.  Using this information, we can calculate the shared
information content of two formats and our measure of
likeliness.

\begin{figure}
  \includegraphics[width=.5\textwidth]{high-level-algorithm.pdf}
  \caption{Schematic diagram for the simple symmetric lens synthesis
  algorithm. The user provides 
   regular expressions \BRegex and \BRegexAlt and a set of examples \Examples{}
    as input. \Expand first converts \BRegex and \BRegexAlt to
    stochastic regular expressions \Regex and \RegexAlt with default
    probabilities. It then finds pairs of stochastic regular expressions
    equivalent to \Regex and \RegexAlt and iteratively proposes them to \GreedySynth.
    \GreedySynth finds a lens typed between the supplied SREs.
    When the algorithm determines it has found a likely lens, it
  returns it.  \ksf{can we change the subscript on the returned lens
  to ``likely'' instead of ``best''? Also, is it the cost that is
  returned or the likeliness }
   }
  \label{fig:high-level-algorithm}
\end{figure}

With simple symmetric lenses and our SRE-based likeliness measure in
hand, we develop a new algorithm for synthesizing likely lenses.
At its core, the algorithm performs a type-directed search between the 
descriptions of the source and target formats, measuring success using
the likeliness measure.
However, regular expressions enjoy a rich algebraic structure, meaning
there are infinitely many regular expressions equivalent to the
original pair, and the lens returned by a type-directed search depends
upon the chosen representation. 
To tame this complexity, we divide the synthesis algorithm into two communicating
search procedures (Figure~\ref{fig:high-level-algorithm}), following
earlier work~\cite{maina+:quotient-synthesis}.
The first algorithm, \Expand, uses rewriting rules to iteratively
propose new pairs of stochastic regular expressions equivalent to the
original pair.
The second algorithm, \GreedySynth, takes a pair of such SREs as input
and uses a greedy, type-directed algoithm to find a simple symmetric
lens between them, returning the lens and its likelihood score
to \Expand.  The algorithm terminates when the likilihood score of the
best lens found so far compared with 
the complexity of newly generated SREs 
suggests that further searching is unlikely to be fruitful.

%%
%%searches for two \emph{related} SREs $S'$ and $T'$ that are
%%\emph{compatible} (a heuristic that estimates the likelihood our second
%%algorithm will succeed). $S'$ and $T'$ are related in the sense that they define
%%the same regular sets as $\BRegex$ and $\BRegexAlt$ (respectively) and the distribution of
%%strings is identical. Indeed, we guarantee these two properties by implementing
%%a search procedure that obeys the star semi-ring axioms of regular expression
%%equivalence, and which we show does not alter the probability distributions of
%%the languages.

%%Given the two compatible SREs, \GreedySynth uses a greedy, type-directed
%%algorithm find a likely lens between them. If algorithm estimates that the lens
%%it has discovered is more likely than any future lens it can possibly find, the
%%search terminates. Otherwise, this lens is stored away (if it is the best found
%%so far) and the search continues with \Expand again.
